---
title: "odejunkfitr"
author: "ojcharles"
date: "19/08/2022"
output: github_document
---

This R package helps with two key problems in viral kinetic modelling.

1 - Asseses whether your ode model is over-parameterised and will fit distributions spuriously.

2 - Fits several viral kinetic models to nlmicr formatted viral load trajectory data, returning which is most parsimonious.


Viral kinetic models often must balance two ideas. It needs sufficient mechanisms to accurately reflect the biology, but also as only one or two compartments can be measured it needs to be sufficiently simple such that key parameters are identifiable. 

This package helps you understand with tangible outcomes where your model lies on that spectrum.
It will fit a model against spurious unique trajectories, if your model can fit them well that could be a problem.
This also will fit your data against several previously published 2, 3 and 4 compartment models that prioritised brevity and identifieability of parameters. Perhaps one of these is better suited.

install with:
`remotes::install_github("ojcharles/does_my_ode_model_fit_junk")`




```{r setup, message = F}
library(ggplot2)
library(nlmixr2)
library(xpose)
library(plotly)
library(odejunkfitr)
odejunkfitr::load_examples() # loads data and models
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

```


#### Define some viral load data, and a current model

```{r load_mod, echo=FALSE}
my_model = function() {
  ini({
    tka <- 0.45
    tcl <- log(c(0, 2.7, 100))
    tv <- 3.45
    eta.ka ~ 0.6
    eta.cl ~ 0.3
    eta.v ~ 0.1
    add.sd <- 0.7
  })
  model({
    ka <- exp(tka + eta.ka)
    cl <- exp(tcl + eta.cl)
    v <- exp(tv + eta.v)
    linCmt() ~ add(add.sd)
  })
}

# example Human Cytomegalovirus untreated viral load trajectories
# Ref: DOI: 10.3390/v13112292
my_data = data2_cmv_untreated
```





### how well does my model fit spurious "junk"?

If the model is capable of fitting all of these trajectories, then you should look in the mirror and contemplate.

#### At the population level

```{r junk_pop, message = F, echo = F}

fit = nlmixr(my_model, data3_junk, list(print=0), est="saem")

# show only the visual predictive check
plots = plot(fit)
plots[28] 
```

#### When fitted individually 

Here we simulate a variable populations around each trajectory then fit in turn to a population of square waves, a population of sine waves etc.

```{r junk_indiv, message =F}

all_fits = data.frame()
for(id in unique(data3_junk$ID)){ # for each type of test data
  
  # spoof a population
  t = data3_junk[data3_junk$ID == id,]
  df2 = t
  for(i in 1:8){
    t1 = t
    t1$ID = i + 1
    t1$DV = t1$DV + rnorm(length(t1$DV), 0, 0.1)
    df2 = rbind(df2,t1)
  }
  
  # get fit
  fit = nlmixr(my_model, df2, list(print=0), est="saem")
  fit$junk_id = id
  all_fits =  rbind(all_fits, fit)
}
  # plot
ggplot(all_fits) +
  geom_point(aes(x = TIME , y = DV)) +
  geom_line(aes(x = TIME, y = IPRED, colour = ID)) +
  theme_classic() +
  labs(subtitle = "Fit of your model to spurious trajectories, per trajectory fit",
       colour = "simulated individual") +
  facet_wrap(~junk_id, scales = "free")

```


 ahh but the dynamics I care about are much faster / shorter!  
 Well then alter the junk data, they're all nicely scaled :)
 
 
```{r junk_indiv_rescale, message =FALSE}
# rescale x axis to be 2 months
data3_junk$TIME = data3_junk$TIME * 5
all_fits = data.frame()
for(id in unique(data3_junk$ID)){ # for each type of test data
  
  # spoof a population
  t = data3_junk[data3_junk$ID == id,]
  df2 = t
  for(i in 1:8){
    t1 = t
    t1$ID = i + 1
    t1$DV = t1$DV + rnorm(length(t1$DV), 0, 0.1)
    df2 = rbind(df2,t1)
  }
  
  # get fit
  fit = nlmixr(my_model, df2, list(print=0), est="saem")
  fit$junk_id = id
  all_fits =  rbind(all_fits, fit)
}
  # plot
ggplot(all_fits) +
  geom_point(aes(x = TIME , y = DV)) +
  geom_line(aes(x = TIME, y = IPRED, colour = ID)) +
  theme_classic() +
  labs(subtitle = "Fit of your model to spurious trajectories, per trajectory fit",
       colour = "simulated individual") +
  facet_wrap(~junk_id, scales = "free") 

data3_junk$TIME = data3_junk$TIME / 5  #rescale back
```




### Fitting multiple viral kinetic models to your data

From the above code you now have a handle for how maleable your model is to various trajectories.
If it has many compartment and many parameters chances are it fitted more than 1 or 2 well.


The following chunks will outline an automated approach to finding the most parsimonious model to fit your viral trajectories. It will retusn information on the AIC (lower is better) as well as the RSE% of estimates for population parameters (above 50% and parameters are considered unidentifiable)

let's plot the example viral load trajectories.

```{r plot_data}
ggplot(my_data) +
  geom_line(aes(x = TIME, y = DV, colour = as.factor(ID))) +
  labs(title = "Human Cytomegalovirus untreated viral load trajectories following BMT", 
       colour = "Event ID", 
       y = "Log10 CMV VL IU/mL", 
       x = "Days since transplant")
```


#### Fit of example model to data

```{r fit_1, message=F}
fit = nlmixr(my_model, my_data, list(print=0), est="saem")
fit$parFixedDf
print(paste0("fit AIC is:", fit$AIC))
```


#### Automated approach to model picking

The following function is quite compute intensive, pre-computed results are shown in this notebook.
- It will run through all the viral kinetic models we ave encoded, currently 9.
- Each model has a number of flaviours, where we have fitted various parameters
- There are currently 24 models
- For each model, we attempt num_fit_attempts , attempts at fitting with random initial pop parameters, and random model seed.
- We found this works well on a machine with at least 16 modern threads

```{r, eval=F}
# This function
df_many_models = odejunkfitr::fit_many_vk_models(vl_data = my_data,
                                                 num_fit_attempts = 5)
df_many_models

```


```{r}
# precomputed AIC comparison
df_many_models = read.csv(system.file("all_vk_models_precomputed.csv", package = "odejunkfitr"))
df_many_models

```


In this example we can see that the models encoded within the odejunkfitr package are more optimal by AIC and result in similarly low maximum RSE% of parameters.


Now you have assessed:
- How well your model fits spurious trajectories, hopefully not too well.
- Whether simple 2,3,4 compartment models are able to reproduce the complex dynamics in your viral load data.


Hopefully this starts you off on a good footing for further model exploration! :)


