---
title: "README"
author: "ojcharles"
date: "19/08/2022"
output: github_document
---

#odejunkfitr

A R package to assess whether your ode model is over-parameterised and will fit any number of distributions. 

In viral kinetic modelling often there are two competing ideas. 1 - I care about uncovering specific mechanistic insights so  my model should represent the biology as accurately as possible. 2 - I can only measure 1 or two compartments, therefore my model should be as simple as possible to allow me to measure the single key parameter I care about.

This package helps you understand with tangible outcomes where your model lies on that spectrum. It does this by fitting against a number of datasets from simple probability distributions to real published viral kinetic data.

install with:
`remotes::install_github("ojcharles/does_my_ode_model_fit_junk")`




## Core usage



```{r setup, echo = F}
## Load libraries
library(ggplot2)
library(nlmixr)
library(xpose)
library(xpose.nlmixr)
library(plotly)
library(odejunkfitr)
odejunkfitr::load_examples()

```


This package takes as input a nlmixr formatted model string.

```{r load_mod, echo=FALSE}
model1_theo_sd <<- "function() {
  ini({
    tka <- 0.45
    tcl <- log(c(0, 2.7, 100))
    tv <- 3.45
    eta.ka ~ 0.6
    eta.cl ~ 0.3
    eta.v ~ 0.1
    add.sd <- 0.7
  })
  model({
    ka <- exp(tka + eta.ka)
    cl <- exp(tcl + eta.cl)
    v <- exp(tv + eta.v)
    linCmt() ~ add(add.sd)
  })
}"

# convert to function
ode_function = ode_string_to_function(model1_theo_sd)

# check model compiles with nlmixr
does_model_compile(ode_function)
```


## fit model to nlmixr typical data



```{r}
fit = nlmixr(ode_function, data1_theo_sd, list(print=0), est="saem") # works

print(paste0("fit AIC is:", fit$AIC))
```


## fit model to lots of spurious data "junk"

this will throw more warnings that a typical fit, and will take longer as this data is so "whack".


```{r}
data_junk = read.csv("inst/junk.csv")
fit = nlmixr(ode_function, data_junk, list(print=0), est="saem") # works

print(paste0("fit AIC is:", fit$AIC))
```

## visualise how well my model fits junk

```{r}
plot(fit)

```


